Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

### 2023-05-05 manuscript/manuscript.qmd
Add more about priors
Create table of summarized predicted probabilities as target

```{r libraries-data}
library(tidyverse)
library(scales)
library(sf)
library(lubridate)
library(tidybayes)
library(marginaleffects)
library(patchwork)
library(colorspace)
library(modelsummary)
library(kableExtra)
library(gt)
library(targets)

@@

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(models_tbl_policies, models_tbl_human_rights))
tar_load(c(derogation_map_data, derogation_count))

@@

fmt_p_inline <- function(x, direction) {
  x <- round(x, 2)

  if (direction == "gt") {
    out <- glue::glue(r"[$p(\Delta > 0) = {x}$]")
  } else {
    out <- glue::glue(r"[$p(\Delta < 0) = {x}$]")
  }

  return(out)
}

policies_pred <- policies_plot_data %>% 
  unnest(pred_data) %>% 
  filter(.width == 0.95) %>% 
  select(nice, year_week_day, derogation_ineffect, .epred, .lower, .upper) %>% 
  group_by(nice, derogation_ineffect) %>% 
  mutate(rank = dense_rank(.epred)) %>% 
  filter(rank == 1 | rank == max(rank)) %>% 
  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
  group_by(rank, derogation_ineffect) %>% 
  mutate(outcome = janitor::make_clean_names(nice)) %>%
  rename(draw = .epred, min = .lower, max = .upper) %>% 
  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0)))) %>% 
  split(~outcome + derogation_ineffect + rank)

policies_mfx <- policies_plot_data %>% 
  unnest(mfx_data) %>% 
  filter(.width == 0.95) %>% 
  select(nice, year_week_day, draw, .lower, .upper, p_gt_0, .width) %>% 
  group_by(nice) %>% 
  mutate(rank = dense_rank(draw)) %>% 
  filter(rank == 1 | rank == max(rank)) %>% 
  mutate(rank = ifelse(rank == 1, "min", "max")) %>% 
  group_by(rank) %>% 
  mutate(outcome = janitor::make_clean_names(nice)) %>%
  rename(min = .lower, max = .upper) %>% 
  mutate(across(c(draw, min, max), list(nice = ~round(. * 100, 0))),
         p_lt_0 = p_gt_0 - 1) %>% 
  mutate(p_gt = fmt_p_inline(p_gt_0, "gt"),
         p_lt = fmt_p_inline(p_lt_0, "lt")) %>% 
  split(~outcome + rank)
```

@@

In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society.

@@

For instance, all countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance. Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021. Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability. The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice` to `r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`). The differences in policy implementation are starker for other types of emergency measures. Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between `r policies_pred$gathering_restrictions.No.min$draw_nice`% and `r policies_pred$gathering_restrictions.No.max$draw_nice`%, representing a `r policies_mfx$gathering_restrictions.min$draw_nice`–`r policies_mfx$gathering_restrictions.max$draw_nice` percentage point difference (`r policies_mfx$gathering_restrictions.max$p_gt`). Derogating states were also far more likely to close public transit, beginning with a `r policies_pred$close_public_transit.Yes.max$draw_nice`% probability in March 2020 and ending in June 2021 with a `r policies_pred$close_public_transit.Yes.min$draw_nice`% probability. Non-derogating states followed a similar downward trend over time, but `r policies_mfx$close_public_transit.min$draw_nice`–`r policies_mfx$close_public_transit.max$draw_nice` percentage points lower than their derogating counterparts (`r policies_mfx$close_public_transit.max$p_gt`). Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to `r policies_pred$movement.Yes.min$draw_nice`% by June 2021. Non-derogating states, on the other hand, began the pandemic with a predicted `r policies_pred$movement.No.max$draw_nice`% probability of limiting movement, dropping to only `r policies_pred$movement.No.min$draw_nice`% by June 2021, representing a `r policies_mfx$movement.min$draw_nice`–`r policies_mfx$movement.max$draw_nice` percentage point difference (`r policies_mfx$movement.max$p_gt`). The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel. Both derogating and non-derogating states had a nearly 100% probability of limiting international movement across all 15 months.

>>>>

```{r libraries-data}
library(tidyverse)
library(scales)
library(sf)
library(lubridate)
library(tidybayes)
library(marginaleffects)
library(patchwork)
library(colorspace)
library(kableExtra)
library(gt)
library(targets)

@@

tar_load(c(policies_plot_data, human_rights_plot_data))
tar_load(c(derogation_map_data, derogation_count))

@@

policies_pred <- policies_table_data$policies_pred_tbl %>% 
  split(~outcome + derogation_ineffect + rank)

policies_mfx <- policies_table_data$policies_mfx_tbl %>% 
  split(~outcome + rank)
```

@@

In all model specifications, we include new and cumulative COVID-19 cases along with indicators for ICCPR derogations or other actions, and V-Dem indices for the respect of rule of law, civil liberties, and civil society. We use Student *t* priors clustered around 0 with a scale parameter of 3, since (1) these model coefficients are measured on the logged odds scale and rarely result in large numbers, and (2) the Student *t* distribution has heavier tails than a standard normal distribution.

@@

For instance, all countries were highly likely to cancel public events in March and April 2020, with a greater than a 95% chance. Over time, countries that did not derogate became less likely to cancel events, with the probability dropping to just above 90% in June 2021. Countries that derogated, however, were essentially guaranteed to regularly cancel events throughout the entire period of study, with a nearly 100% probability. The difference in predicted probabilities between the two types of countries ranged from `r policies_mfx$cancel_public_events.min$draw_nice`–`r policies_mfx$cancel_public_events.max$draw_nice` percentage points (`r policies_mfx$cancel_public_events.max$p_gt`) between March 2020 and June 2021. The differences in policy implementation are starker for other types of emergency measures. Derogating states had a nearly 100% probability of imposing restrictions on gathering across all 15 months, while non-derogating states ranged between `r policies_pred$gathering_restrictions.No.min$draw_nice`% and `r policies_pred$gathering_restrictions.No.max$draw_nice`%, representing a `r policies_mfx$gathering_restrictions.min$draw_nice`–`r policies_mfx$gathering_restrictions.max$draw_nice` percentage point difference (`r policies_mfx$gathering_restrictions.max$p_gt`). Derogating states were also far more likely to close public transit, beginning with a `r policies_pred$close_public_transit.Yes.max$draw_nice`% probability in March 2020 and ending in June 2021 with a `r policies_pred$close_public_transit.Yes.min$draw_nice`% probability. Non-derogating states followed a similar downward trend over time, but `r policies_mfx$close_public_transit.min$draw_nice`–`r policies_mfx$close_public_transit.max$draw_nice` percentage points lower than their derogating counterparts (`r policies_mfx$close_public_transit.max$p_gt`). Restrictions on internal movement were practically guaranteed at the beginning of the pandemic for derogating states, with a nearly 100% probability, dropping to `r policies_pred$movement.Yes.min$draw_nice`% by June 2021. Non-derogating states, on the other hand, began the pandemic with a predicted `r policies_pred$movement.No.max$draw_nice`% probability of limiting movement, dropping to only `r policies_pred$movement.No.min$draw_nice`% by June 2021, representing a `r policies_mfx$movement.min$draw_nice`–`r policies_mfx$movement.max$draw_nice` percentage point difference (`r policies_mfx$movement.max$p_gt`). The only policy we observed where there was no sizable difference between the two types of states is restrictions on international travel. Both derogating and non-derogating states had a nearly 100% probability of limiting international movement across all 15 months.
